---
title: "Covid91 vaccine study Final2023"
author: "lana"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ The covid-921 vaccine does not exhibit a significant difference in effectiveness between males 

$H_A$ the covid-921 vaccine does exhibit a significant difference in effectiveness between males 

```{r}
library(dplyr)
males = filter(FinalData, FinalData$sex=="M")
```

##Methods 

Males is categorical and covid-921 is explanatory, therefore we use cat~cat. 

##Graphical Results

```{r}
barchartGC(~infected + treatment,data=males)
barchartGC(~infected + treatment,data=males, type="percent")
```

this bar graph shows males who has covid and took the vaccine did better than the males who didn't take the vaccine the males who dont have covid who took the vaccine stayed fairly the same to those who didn't take the vaccine. 

```{r}
table1 <- xtabs(~infected + treatment, data=males)
rowPerc(table1)
colPerc(table1)
```

this table shows the males who have covid that took the vaccine, 30% still had covid and people who took the placebo, 69% still have covid. the males without covid who took the vaccine, 49%l got covid and the males who took the placebo, 50% got covid. 


## Inferntal Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the p value is 0.0000000000001098 we reject the null hypothesis. 

## Females

$H_0$ the covid-921 vaccine does not exhibit a significant defference in effectiveness between female 

$H_A$ the covid-921 vaccine does not exhibit a significant defference in effectiveness between female 

```{r}
library(dplyr)
female = filter(FinalData, FinalData$sex=="F")
```


## Methods 

Males is categorical and covid-921 is explanatory, therefore we use cat~cat.

## Graphical Results 

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```


## LGBTQ

$H_0$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between lgbtq 

$H_A$ the covid-921 vaccine does exhibit a significant difference in effectiveness between lgbtq

```{r}
library(dplyr)
lgbtq = filter(FinalData, FinalData$LGBTQ=="gay")
```

##Methods

Males is categorical and covid-921 is explanatory, therefore we use cat~cat

##Graphical Results 

```{r}
barchartGC(~infected + treatment,data=lgbtq)
barchartGC(~infected + treatment,data=lgbtq, type="percent")
```


## Druggies

$H_0$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between drugies 

$H_A$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between drugies 

```{r}
library(dplyr)
druggies = filter(FinalData, FinalData$DrugUser== "Yes")
```

##Graphical Results 

```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")
```

# Overall Results and Conclusions


The covid-921 vaccine does not exhibit a significant difference in effectiveness between males. Males is categorical and covid-921 is explanatory, therefore we use cat~cat. The bar graph shows males who had covid and took the vaccine did better than the males who didn't take the vaccine. The males who didn't have covid who took the vaccine stayed fairly the same to those who didn't take the vaccine. This table shows the males who have covid that took the vaccine, 30% still had covid and people who took the placebo, 69% still have covid. of the males without covid who took the vaccine, 49% got covid and the males who took the placebo, 50% got covid. the p value is 0.0000000000001098 we reject the null hypothesis. The covid-921 vaccine does not exhibit a significant difference in effectiveness between females. The covid-921 vaccine does not exhibit a significant difference in effectiveness between lgbtq. The covid-921 vaccine does not exhibit a significant difference in effectiveness between druggies. 




